---
title: "bootstrap"
author: "Xiaoni Xu"
date: "2024-11-14"
output: html_document
---
```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

## Bootstrapping in SLR
```{r}
n_samp = 250

sim_df_const = 
  tibble(
    x = rnorm(n_samp, 1, 1),
    error = rnorm(n_samp, 0, 1),
    y = 2 + 3 * x + error
  )

sim_df_nonconst = sim_df_const |> 
  mutate(
  error = error * .75 * x,
  y = 2 + 3 * x + error
)
```

```{r}
sim_df = 
  bind_rows(const = sim_df_const, nonconst = sim_df_nonconst, .id = "data_source") 

sim_df |> 
  ggplot(aes(x = x, y = y)) + 
  geom_point(alpha = .5) +
  stat_smooth(method = "lm") +
  facet_grid(~data_source) 
```

Look at regression results

```{r}
lm(y ~ x, data = sim_df_const) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```

```{r}
lm(y ~ x, data = sim_df_nonconst) |> 
  broom::tidy() |> 
  knitr::kable(digits = 3)
```

## Drawing one bootstrap sample

```{r}
boot_sample = function(df) {
  
  boot_df =
    sample_frac(df, replace = TRUE) %>% 
    arrange(x)
  
  return(boot_df)
}
```

Try running this

```{r}
sim_df_nonconst %>% 
  boot_sample() %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point(alpha = .5) +
  stat_smooth(method = "lm")
```

## Bootstrap a lot

```{r}
boot_straps = 
  tibble(
    strap_number = 1:10
  ) %>% 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = sim_df_nonconst)),
    models = map(strap_sample, \(df) lm(y ~ x, data = df))
  )
```

## Analyzing bootstrap samples

```{r}
bootstrap_results = 
  boot_straps |> 
  mutate(
    models = map(strap_sample, \(df) lm(y ~ x, data = df) ),
    results = map(models, broom::tidy)) |> 
  select(-strap_sample, -models) |> 
  unnest(results) 

bootstrap_results |> 
  group_by(term) |> 
  summarize(boot_se = sd(estimate)) |> 
  knitr::kable(digits = 3)
```

I can also use the estimates across bootstrap samples to construct a confidence interval. For a 95% CI, we might try to exclude the lower and upper 2.5% of the distribution of parameter estimates across “repeated” samples. The code below will do that.

```{r}
bootstrap_results |> 
  group_by(term) |> 
  summarize(
    ci_lower = quantile(estimate, 0.025), 
    ci_upper = quantile(estimate, 0.975))
```

For a simple linear regression, we can show the fitted lines for each bootstrap sample to build intuition for these results.

```{r}
boot_straps |> 
  unnest(strap_sample) |> 
  ggplot(aes(x = x, y = y)) + 
  geom_line(aes(group = strap_number), stat = "smooth", method = "lm", se = FALSE, alpha = .1, color = "blue") +
  geom_point(data = sim_df_nonconst, alpha = .5)
```

## Bootstrap function

Bootstrapping is common enough that it’s been automated, to some degree, in the `modelr::boostrap` function

```{r}
boot_straps = 
  sim_df_nonconst |> 
  modelr::bootstrap(n = 1000)

boot_straps |> pull(strap) |> nth(1)
```

```{r}
boot_straps |> pull(strap) |> nth(1) |> as_tibble()
```

Let’s repeat our analysis pipeline using the `bootstrap` function instead of our own process for drawing samples with replacement.

```{r}
sim_df_nonconst |> 
  modelr::bootstrap(n = 1000) |> 
  mutate(
    models = map(strap, \(df) lm(y ~ x, data = df) ),
    results = map(models, broom::tidy)) |> 
  select(-strap, -models) |> 
  unnest(results) |> 
  group_by(term) |> 
  summarize(boot_se = sd(estimate))
```

To bootstrap the dataset with constant error variance, we only have to change the input dataframe
```{r}
sim_df_const |> 
  modelr::bootstrap(n = 1000) |> 
  mutate(
    models = map(strap, \(df) lm(y ~ x, data = df)),
    results = map(models, broom::tidy)) |> 
  select(-strap, -models) |> 
  unnest(results) |> 
  group_by(term) |> 
  summarize(boot_se = sd(estimate))
```

## Airbnb data

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  drop_na(price, stars) |> 
  select(price, stars, borough, neighborhood, room_type)
```

```{r}
nyc_airbnb |> 
  ggplot(aes(x = stars, y = price, color = room_type)) + 
  geom_point() 
```

```{r}
nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  modelr::bootstrap(n = 1000) |> 
  mutate(
    models = map(strap, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)) |> 
  select(results) |> 
  unnest(results) |> 
  filter(term == "stars") |> 
  ggplot(aes(x = estimate)) + geom_density()
```



