---
title: "linear_models"
author: "Xiaoni Xu"
date: "2024-11-07"
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r, message = FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

Do some data cleaning

```{r}
data("nyc_airbnb")

nyc_airbnb =
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood # AMERICA FIRST
  ) %>% 
  filter(
    borough != "Staten Island"
  ) %>% 
  select(price, stars, borough, neighborhood, room_type) %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  )
```

### Fit some models

Start relatively simple
```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

summary(fit) # a lot of stuff comes out

names(summary(fit))

coef(fit)

broom::tidy(fit)

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)

fit %>% 
  broom::glance()
```

Try something a bit more complex
```{r}
fit =
  lm(price ~ stars + borough, data = nyc_airbnb)
# take all boroughs, take first one as reference, and make comparisons with that one. continue with the rest of the boroughs.

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) %>% 
  knitr::kable(digits = 3)
```

## Some diagnostics

Backtrack to some EDA

```{r}
nyc_airbnb %>% 
  ggplot(aes(x = stars, y = price)) +
  geom_point() + 
  stat_smooth(method = "lm")
```



Most diagnostics use residuals

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = resid)) +
  geom_histogram()
```

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin() +
  ylim(-100, 100)
```

Residuals against stars

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```

Residuals against fitted values
```{r}
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  modelr::add_predictions(fit) %>% 
  ggplot(aes(x = pred, y = resid)) +
  geom_point()
```

## Hypothesis testing

For single coefficients, look at the table
```{r}
fit %>% 
  broom::tidy()

fit_null = lm(price ~ stars + borough, data = nyc_airbnb)

fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)

```

Look at both
```{r}
fit_null %>% 
  broom::tidy()

fit_alt %>% 
  broom::tidy()

anova(fit_null, fit_alt) %>% 
  broom::tidy()
```

## Do effects differ across boroughs

First, use a lot of interactions

```{r}
nyc_airbnb %>% 
  lm(price ~ stars * borough + room_type * borough, data = .) %>% # use "." instead of "_" after "data = ".
  broom::tidy()
```

Could just fit separate models.

```{r}
nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy()

nyc_airbnb %>% 
  filter(borough == "Brooklyn") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy()
```

Get fancy and use list columns

```{r}
nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
  ) %>% 
  select(borough, results) %>% 
  unnest(results) %>% 
  select(borough, term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```

What about room type across Manhattan neighborhoods?

```{r}
nyc_airbnb %>% 
  filter(borough == "Manhattan",
         neighborhood == "Chinatown") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy()

nyc_airbnb %>% 
  filter(borough == "Manhattan",
         neighborhood == "Chelsea") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy()
```

Do this for all neighborhoods

```{r}
manhattan_neighborhood_fit_df =
  nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  nest(data = -(borough:neighborhood)) %>% 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
  ) %>% 
  select(neighborhood, results) %>% 
  unnest(results)
```

Look at the effect of room type
```{r}
manhattan_neighborhood_fit_df %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = term, y = estimate)) +
  geom_boxplot()
```

```{r}
manhattan_neighborhood_fit_df %>% 
  filter(str_detect(term, "room_type")) %>% 
  filter(estimate > 0)

nyc_airbnb %>% 
  filter(
    neighborhood == "NoHo",
    room_type == "Shared room"
  )

# the outlier in the shared room boxplot comes from NoHo only having one shared room listing that is very expensive.
```

